Measurement of total hadronic differential cross sections in the LArIAT experiment

### 2018-04-10 Chapter-1/1-Theory.tex
s

\subsection{Make up of Neutrino Interactions}
All neutrino experiments involving the detection of single neutrinos are concerned with neutrino interactions (and neutrino cross sections) on nuclei. 
Given the invisible nature of the neutrino, characterizing the products of its interaction is the only method to a) assess the neutrino existence, b) detect its flavor in case of a charge current interaction and c) eventually reconstruct its energy. The complexity of the nuclear structure complicate this task. Nuclear effects modify the kinematic and final state of the neutrino interaction products and play a key role in modern neutrino experiments that use fairly heavy nuclear targets.  Data on neutrino scattering off many different nuclei are available for both charged current (CC) and neutral current (NC) channels, as summarized here \cite{RevModPhys.84.1307}.

Historically, neutrino interactions with the nucleus are divided into three categories as a function of increasing neutrino energy: quasi elastic, resonant and deep inelastic scattering.
All current and forthcoming oscillation experiments live in
the 0.1-10 GeV transition region, which encompasses the energy where the quasi-elastic neutrino-nucleus interaction transitions into resonance scattering and the energy  resonance scattering  transitions in to deep inelastic scattering. 
\textcolor{red}{PLOT}

%%%%%%%%%%%%%%%%%%%%%%%%  
\textbf{Quasi-elastic scattering}
Quasi-elastic (QE) scattering is the dominant neutrino interaction for neutrino energies less than ? 1 GeV and represents a large fraction of the signal samples in many neutrino oscillation experiments. Historically, neutrino (antineutrino) quasi-elastic scattering refers to the process $\nu_\mu n \rightarrow \mu^- p$ and ($\bar\nu_\mu p \rightarrow \mu^+ n$) where a charged lepton and single nucleon are ejected in the elastic interaction of a neutrino (or antineutrino) with a nucleon in the target material. This is the final state one would strictly observe, for example, in scattering off of a free nucleon target. %% ADD DIFFERENCES WITH HEAVIER TARGETS
In many of these initial measurements of the neutrino QE cross section, bubble chamber experiments employed light targets (H2 or D2) and required both the detection of the final state muon and single nucleon; thus the final state was clear and elastic kinematic conditions could be verified. The situation is more complicated, of course, for heavier nuclear targets. In this case, nuclear effects can impact the size and shape of the cross section as well as the final state kinematics and topology. Due to intra-nuclear hadron rescattering and the possible effects of correlations between target nucleons, additional nucleons may be ejected in the final state; hence, a QE interaction on a nuclear target does not always imply the ejection of a single nucleon. 
One therefore needs to take some care in defining what one means by neutrino QE scattering when scattering off targets heavier than H2 or D2.

Fig. 49.2 displays the current status of existing measurements of $\nu_\mu$ and $\bar\nu_\mu$ QE scattering cross sections as a function of neutrino energy. In this plot, and all others in this review, the prediction from a representative neutrino event generator (NUANCE) [46] provides a theoretical comparator. MiniBooNE XS enanchement Such an enhancement is believed to be due to the presence of correlations between target nucleons in the nucleus. As a result, the impact of such nuclear effects on neutrino QE scattering has recently been the subject of intense experimental and theoretical scrutiny with potential implications on event rates, nucleon emission, neutrino energy reconstruction, and neutrino/antineutrino ratios. 
Experiments such as ArgoNeuT have begun to provide the first measurements of proton multiplicities in neutrino-argon QE scattering [4,48], a critical ingredient in understanding the hadronic side of these interactions and final state effects.

%Note that modern experiments have recently opted to report QE cross sections as a function of final state muon or proton kinematics [17,18,48]. Such distributions are more difficult to compare between experiments but are much less model-dependent and provide more stringent tests of the theory than cross sections as a function of neutrino energy (E?) or 4-momentum transfer (Q2).

In addition to such charged current investigations, measurements of the neutral current counterpart of this channel have also been performed. The most recent NC elastic scattering cross section measurements include those from BNL E734 [62], MiniBooNE [20,21,22], and T2K [40].

\textbf{Pion Production}
In addition to such elastic processes, neutrinos can also inelastically scatter producing a nucleon excited state ($\Delta$,N*). Such baryonic resonances quickly decay, most often to a nucleon and single-pion final state. Historically, experiments have measured various exclusive final states associated with these reactions, the majority of which have been on hydrogen and deuterium targets [1].
In addition to such resonance production processes, neutrinos can also coherently scatter off of the entire nucleus and produce a distinctly forward-scattered single pion final state. Both CC %(??A???A?+,??A??+A?? ) 
and NC %(??A???A?0,??A???A?0) 
processes are possible in this case. Even though the level of coherent pion production is small compared to resonant processes, observations exist across a broad energy range and on multiple nuclear targets [64]. More recently, several modern neutrino experiments have measured coherent pion production cross sections including ArgoNeuT [5], K2K [8], MINERvA [15], MiniBooNE [27], MINOS [65], NOMAD [32], and SciBooNE [34,36].
As with QE scattering, a new appreciation for the significance of nuclear effects has surfaced in pion production channels, again due to the use of heavy nuclear targets in modern neutrino experiments. Many experiments have been careful to report cross sections for various detected final states, thereby not correcting for large and uncertain nuclear effects (e.g., pion re-scattering, charge exchange, and absorption) which can introduce unwanted sources of uncertainty and model dependence. Recent measurements of single-pion cross sections, as published by K2K [9,10], MiniBooNE [24], and SciBooNE [35], take the form of ratios with respect to QE or CC inclusive scattering samples. Providing the most comprehensive survey of neutrino single-pion production to date, MiniBooNE has recently published a total of 16 single- and double-differential cross sections for both the final state muon (in the case of CC scattering) and pion in these interactions; thus, providing the first measurements of these distributions (Fig. 49.3) [23?26]. MINERvA has recently produced similar kinematic measurements at higher neutrino energies (Fig. 49.4) [16,66]. Regardless of the interaction channel, such differential cross section measurements (in terms of observed final state particle kinematics) are now preferred for their reduced model dependence and for the additional kinematic information they provide. Such a new direction has been the focus of modern measurements as opposed to the reporting of more model-dependent, historical cross sections as a function of E? or Q2. Together with similar results for other interaction channels, a better understanding and modeling of nuclear effects will be possible moving forward.It should be noted that baryonic resonances can also decay to multi-pion, other mesonic (K, ?, ?, etc.), and even photon final states. Experimental results for these channels are typically sparse or non-existent [1]; however, photon production processes can be an important background for ?? ? ?e appearance searches and thus have become the focus of some recent experimental investigations; for example, in NOMAD [67].

\textbf{Inclusive Scattering -- DIS}
Over the years, many experiments have measured the total inclusive cross section ($\nu_\mu N \rightarrow \mu^- X$) and antineutrino ($\bar\nu_\mu N \rightarrow \mu^+ X$) scattering off nucleons covering a broad range of neutrino energies. As can be seen in Fig. 49.1, the inclusive cross section approaches a linear dependence on neutrino energy. Such behavior is expected for point-like scattering of neutrinos from quarks, an assumption which breaks down at lower energies. To provide a more complete picture, differential cross sections for such inclusive scattering processes have been reported ? these include measurements on iron from NuTeV [42] and, more recently, at lower neutrino energies on argon from ArgoNeuT [2,3] and on carbon from T2K [37]. MINERvA has also provided new measurements of the ratios of the muon neutrino CC inclusive scattering cross section on a variety of nuclear targets such as lead, iron, and carbon [11]. At high energy, the inclusive cross section is dominated by deep inelastic scattering (DIS). Several high energy neutrino experiments have measured the DIS cross sections for specific final states, for example opposite-sign dimuon production. The most recent dimuon cross section measurements include those from CHORUS [43], NOMAD [44], and NuTeV [45]. At lower neutrino energies, the inclusive cross section is an additionally complex combination of quasi-elastic scattering

%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

\subsection{Make up of Neutrino Interactions}
All neutrino experiments involving the detection of single neutrinos are concerned with neutrino interactions (and neutrino cross sections) on nuclei. 
Given the invisible nature of the neutrino, characterizing the products of its interaction is the only method to a) assess the neutrino presence, b) detect its flavor in case of a charge current interaction and c) eventually reconstruct its energy.

Historically, neutrino interactions with the nucleus in the GeV region are divided into three categories as a function of increasing neutrino energy: quasi elastic (QE), resonant and deep inelastic scattering (DIS). All current and forthcoming oscillation experiments live in the 0.1-10 GeV transition region, which encompasses the energy where the QE neutrino-nucleus interaction transitions into resonant scattering and the energy  where resonance scattering  transitions in to DIS. 
 Neutrino and antineutrino QE charge current scattering refers to the process $\nu_\mu n \rightarrow \mu^- p$ and $\bar\nu_\mu p \rightarrow \mu^+ n$ where a charged lepton and single nucleon are ejected in the elastic interaction and the target nucleus remains at ground state. %of a neutrino (or antineutrino) with a nucleon in the target material.  
 Resonant scattering refers to an inelastic collision producing a nucleon excited state ($\Delta$,N*); the excited resonance quickly decays, most often to a nucleon and single-pion final state. DIS refers to the head-on collision between the neutrino and a parton inside the nucleon, producing hadronization and subsequent  abundant production of mesons and nucleons.  In addition to such interactions between the neutrino and a single component of the nucleus, neutrinos can also interact with the nucleus as a whole, albeit more rarely, a well documented process called coherent meson production scattering \cite{Vilain1993}; the signature of such process is the production of a distinctly forward-scattered single meson final state, most often a pion.  This simple picture of neutrino interactions works rather well for scattering 
off of light targets, such as the H$_2$ and D$_2$ of bubble chamber experiments \cite{RevModPhys.84.1307}, but the complexity of the nuclear structure for heavier nuclei such as argon complicates this model.

As we will discuss in Chapter \ref{ch:2}, the properties of argon make it a good candidate for interacting medium in neutrino experiments; in particular the density of its interaction centers  augments the yield of neutrino interactions and allows for relatively compact detectors. Though, the choice of a relatively heavy nuclear target comes at the cost of enhancing  nuclear effects which modify the kinematic and final state of the neutrino interaction products.

Nuclear effects can potentially affect the neutrino event rates, nucleon emission, neutrino energy reconstruction, and neutrino/antineutrino ratios, carrying deep implications for oscillation experiments. 
Even in the case of ``simple" QE scattering  nuclear effects can impact the size and shape of the cross section as well as the final state kinematics and topology. Intra-nuclear hadron rescattering and correlation effects between the target nucleons can cause the ejection of additional nucleons in the final state. In case of resonant and DIS scattering, the hadronic interactions of meson and nucleons produced in the decay of the resonance or during hadronization  complicate this picture even more.
A large source of uncertainty in modeling  nuclear effects in neutrino interactions come from mesons interactions (and re-interactions) in the nucleus, e.g., pion re-scattering, charge exchange, and absorption.

A renewed interest for neutrino cross section measurements surged in recent year, along with lively discussion on the data reporting; the historical method of reporting the neutrino cross section as a function of the neutrino energy or momentum transferred shakes under the weight of its dependency to the chosen nuclear model.   On one hand, correcting for nuclear effects in neutrino interaction can introduce unwanted sources of uncertainty and model dependence due to the mis-modeling of the meson interactions. On the other, avoiding this correction however makes a comparison between neutrino interactions on different target nuclei extremely difficult.

Data on neutrino scattering off many different nuclei are available for both charged current (CC) and neutral current (NC) channels, as summarized here \cite{RevModPhys.84.1307}. A summary of is reported in Figure \textcolor{red}{PLOT} where the (NUANCE) [46] event generator provides a theoretical comparator.
